---
title: "a5-data-report"
author: "Alan Li"
date: "11/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(knitr)
library(tidyr)
library(leaflet)
library(ggplot2)
```

```{r include=FALSE}
source("analysis.R")
```

## Summary Information

Based on the [shootings 2018 dataset](http://www.shootingtracker.com), there 
are `r gun_shooting_occurencies` shootings occurred in the year of 2018. Also, `r lives_lost` lives 
were lost during those `r gun_shooting_occurencies` shootings. Moreover, city that was most impacted 
by shootings is `r most_impact_city`. The state that was most impacted by shootings 
is `r most_impact_state`. And the date which is most impacted by shootings is `r most_horrible_day`.

## Summary Table
`r state_analysis_table`

* In this table, I grouped the data by states and calculate the total death number by each state.
Then I sort the number by a decreasing order. It turns out California as the most dangerous state
when it comes to gunshooting and Utah is the safest.

## Description of a Particular Incident

The most horrible gun shooting recorded would be the one located at `r city` with the most people 
killed. The incident occured at `r date_of_event` and it happened at `r address`, `r city`, 
`r state`. In this shooting `r number_killed` souls were lost and the total number of people injured
is `r number_injured`. The killing spree is the deadliest high school shooting in United States 
history, surpassing the Columbine High School massacre that killed 15, including the perpetrators, 
in Colorado in April 1999. 

Link to Additional Information: 
<https://www.history.com/this-day-in-history/parkland-marjory-stoneman-douglas-school-shooting>

## Interactive Map

`r map`

Each marker represent a gun shooting event. The radius of the marker represent the total casualties 
of the shooting. As shown on the map, east coast has significantly more shootings than west coast. 
The shootings tends to happen at major cities like New York and Chicago.

## Plot of your choice
```{r mass_shootings_data, echo=FALSE, warning=FALSE}
library(ggplot2)
ggplot(data = gun_shooting, aes(state, num_killed)) + geom_bar(stat="identity") +
theme(axis.text.x = element_text(size  = 10, angle = 45, hjust = 1, vjust = 1))
```

## Question
* Which state(s) has significantly higer deaths compared to others?

## Insights
*California and Florida has nearly 50 deaths during the whole year of 2018. Which is significantly
more than other states
*Connecticut, Massachusetts and Utah has not even one single death by shooting in the whole year of
2018

